A survey of rigour and reporting standards for gene set enrichment tests

### 2021-08-11 manuscript.Rmd
minor update to ms

6. If ORA is used, it is vital to use a background list consisting of all detected genes.
Do not use tools that cannot accept a background list.

7. Criteria and parameters matter.
If performing ORA, then the inclusion thresholds need to be mentioned.
Moreover it needs to be mentioned whether up- and down-regulated genes are combined or separated into different tests.
If using GSEA, parameters around gene weighting and type of test need to be disclosed (eg: permuting sample labels or gene labels).

@@

8. Preference FCS and pathway topology tools over ORA tools.
ORA tools have a lower sensitivity. 
In practice, they are rarely used properly.
FCS tests distinguish between up- and down-regulated gene sets by default.

9. Include the gene profile data (including any background lists) in the supplementary data in TSV or CSV formats.
Excel spreadsheets are not recommended for genomics work (Ziemann et al, 2016)

10. Scripted analysis workflows are preferred over analysis involving graphical user interfaces because they can attain a higher degree of computational reproducibility, so consider investing in upgrading your analysis.
Similarly, free and open source software tools are preferred over proprietary software.

11. If using a scripted analysis, provide code by depositing it to a repository like GitHub or Zenodo.
Link the data to the code, so anyone can download and reproduce your analysis (Peng 2011).

>>>>

6. If ORA is used, it is vital to use a background list consisting of all detected genes, and report this in the methods section.
Avoid tools that cannot accept a background list.

7. Criteria and parameters matter.
If performing ORA, then the inclusion thresholds need to be mentioned.
If using GSEA or another FCS tool, parameters around ranking, gene weighting and type of test need to be disclosed (eg: permuting sample labels or gene labels).

8. If using ORA, it needs to be mentioned whether up- and down-regulated genes are separated into different tests as recommended by Hong et al (2014) or combined.

@@

9. Preference FCS and pathway topology tools over ORA tools.
ORA tools have a lower sensitivity. 
In practice, they are rarely used properly.
FCS tests distinguish between up- and down-regulated gene sets by default.

10. Include the gene profile data (including any background lists) in the supplementary data in TSV or CSV formats.
Excel spreadsheets are not recommended for genomics work (Ziemann et al, 2016)

11. Scripted analysis workflows are preferred over analysis involving graphical user interfaces because they can attain a higher degree of computational reproducibility, so consider investing in upgrading your analysis.
Similarly, free and open source software tools are preferred over proprietary software.

12. If using a scripted analysis, provide code by depositing it to a repository like GitHub or Zenodo.
Link the data to the code, so anyone can download and reproduce your analysis (Peng 2011).
